---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Cailey Fay"
date: "10.14.25"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
library(lemon)
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1
```{r exercise1}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)

lq_ak <- laquinta %>%
  filter(state == "AK")

nrow(lq_ak)
```

There are `r nrow(dn_ak)` Denny's and `r nrow(lq_ak)` Laquinta's in Alaska. 

### Exercise 2

There are 6 Pairings.
```{r exercise2}
pairings <- nrow(dn_ak)*nrow(lq_ak)
```
### Exercise 3 and 4
``` {r exercise3 and 4}

dn_lq_ak <- full_join(dn_ak, lq_ak,
  by = "state")
dn_lq_ak

```

There are `r nrow(dn_lq_ak)` observations, each representing a denny's and laquinta pairing in Alaska. 

### Exercise 5

We use the mutate function! 
``` {r exercise 5}

haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 <- long1 * pi / 180
  lat1 <- lat1 * pi / 180
  long2 <- long2 * pi / 180
  lat2 <- lat2 * pi / 180

  R <- 6371 # Earth mean radius in km

  a <- sin((lat2 - lat1) / 2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1) / 2)^2
  d <- R * 2 * asin(sqrt(a))

  return(round(d, round)) # distance in km
}
```
...

### Exercise 6
```{r exercise6}

dn_lq_ak <- dn_lq_ak %>%
  mutate(distance = haversine(longitude.x,latitude.x,longitude.y,latitude.y))

```

...
### Exercise 7
```{r exercise7}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
```

### Exercise 8
```{r exercise8}

#From Lab 4, visualization of Dennys and Laquintas throughout the United States.
dn <- dennys %>%
  mutate(establishment = "Denny's")

lq <- laquinta %>%
  mutate(establishment = "La Quinta")

dn_lq <- bind_rows(dn, lq)

ggplot(dn_lq, mapping = aes(
  x = longitude,
  y = latitude,
  color = establishment
)) +
  geom_point() +
  labs(title = "Denny's and Laquintas in the United States",
       x = "Longitude",
       y = "Latitude")

# Looking at the map of Alaska only. 
AKonly <- dn_lq %>%
  filter(state == "AK")
ggplot(AKonly, mapping = aes(
  x = longitude,
  y = latitude,
  color = establishment
)) +
  geom_point()  +
  labs(title = "Denny's and Laquintas in Alaska",
       x = "Longitude",
       y = "Latitude")

```

Every Denny's has a Laquinta within six miles, according to the minimum distance variable created in exercise 7. The average distance is `r mean(dn_lq_ak_mindist$closest)` with a standard deviation of `r sd(dn_lq_ak_mindist$closest)`. 

### Exercise 9
```{r exercise9}
dn_nc <- dennys %>%
  filter(state == "NC")

lq_nc <- laquinta %>%
  filter(state == "NC")

dn_lq_nc <- full_join(dn_nc, lq_nc,
  by = "state")

dn_lq_nc <- dn_lq_nc %>%
  mutate(distance = haversine(longitude.x,latitude.x,longitude.y,latitude.y))

dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

dn_lq_nc_mindist %>%
  ggplot(aes(x=closest)) +
  geom_histogram(binwidth = 25) +
  labs(title ="Distribution of Minimum Distances Between Denny's and Laquinta's",
       subtitle = "In North Carolina",
       x= "Miles",
       y = "Frequency")

```

On average, there is an Laquinta within `r mean(dn_lq_nc_mindist$closest)` miles of a dennys, with a standard deviation of `r sd(dn_lq_nc_mindist$closest)`. This distribution is much more spread out than alaska, with some "nearest pairings" close by, and others really far apart. The farthest Denny's to a Laquinta is 187 miles. 

### Exercise 10
```{r}
dn_tx <- dennys %>%
  filter(state == "TX")

lq_tx <- laquinta %>%
  filter(state == "TX")

dn_lq_tx <- full_join(dn_tx, lq_tx,
  by = "state")

dn_lq_tx <- dn_lq_tx %>%
  mutate(distance = haversine(longitude.x,latitude.x,longitude.y,latitude.y))

dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

dn_lq_tx_mindist %>%
  ggplot(aes(x=closest)) +
  geom_density() +
   labs(title ="Distribution of Minimum Distances Between Denny's and Laquinta's",
       subtitle = "In Texas",
       x= "Miles",
       y = "Frequency")

```

On average, there is an Laquinta within `r mean(dn_lq_tx_mindist$closest)` miles of a dennys, with a standard deviation of `r sd(dn_lq_tx_mindist$closest)`. This distribution is much more condensed than North Carolina, with some "nearest pairings" close by, and others really far apart. The farthest Denny's to a Laquinta is 60 miles, but many locations are within 1/2 a mile. 

### Exercise 11
```{r exercise 11}

dn_fl <- dennys %>%
  filter(state == "FL")

lq_fl <- laquinta %>%
  filter(state == "FL")

dn_lq_fl <- full_join(dn_fl, lq_fl,
  by = "state")

dn_lq_fl <- dn_lq_fl %>%
  mutate(distance = haversine(longitude.x,latitude.x,longitude.y,latitude.y))

dn_lq_fl_mindist <- dn_lq_fl %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

dn_lq_fl_mindist %>%
  ggplot(aes(x=closest)) +
  geom_density() +
   labs(title ="Distribution of Minimum Distances Between Denny's and Laquinta's",
       subtitle = "In Florida",
       x= "Miles",
       y = "Frequency")

```

On average, there is an Laquinta within `r mean(dn_lq_fl_mindist$closest)` miles of a dennys, with a standard deviation of `r sd(dn_lq_fl_mindist$closest)`. This distribution is much more condensed than North Carolina, with some "nearest pairings" close by, and others really far apart. The farthest Denny's to a Laquinta is 176 miles, but many locations are within a mile or two. The 176 mile one is very much an outlier, as is the one that is 90 miles away. 

### Exercise 12 

I think the joke is most true for Texas, because the proportion of Denny's within a mile from a Laquinta is greatest, adjusting for the total number of Dennys in the state. 56/200 are within a mile in Texas, whereas 13/139 are within a mile in Florida. 

```{r}
dn_lq_ak_mindist %>%
  filter(closest < 1) %>%
  count()
#none are within 0 miles in Alaska 

dn_lq_nc_mindist %>%
  filter(closest < 1) %>%
  count()
#none are within 0 miles in North Carolina 

dn_lq_tx_mindist %>%
  filter(closest < 1) %>%
  count()
#there are 56 within 0 miles in Texas 
#out of 200

dn_lq_fl_mindist %>%
  filter(closest < 1) %>%
  count()
#There are 16 within 0 miles in Florida 
# out of 139

```


### Trying to recreate those plots: Not quite recreated but as much time as I am willing to invest. 
```{r}
dn_lq_dist <- full_join(dn, lq, by = "state")

dn_lq_dist <- dn_lq_dist %>%
  mutate(distance = haversine(longitude.x,latitude.x,longitude.y,latitude.y))

dn_lq_mindist <- dn_lq_dist %>%
  group_by(address.x) %>%
  mutate(closest = min(distance))

dn_lq_mindist %>%
  filter(state %in% c("NC","ND","NE","NH","NJ","NM","NV","NY")) %>%
  ggplot(aes(x=closest, fill=state)) +
  geom_density() +
  facet_grid(state ~ ., switch = "y", scales = "free_y") +
  theme_minimal() +
  theme(axis.text.y = element_blank()) +
  labs(title = "Minimum Distances Between Denny's and Laquinta's By State",
       x = "Nearest Location (In Miles)",
       y = "Density",
       legend = "State")
 
```
